---
title: "Assignment #1"
author : "Tianqi Wang"
date: "July 16, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,tidy.opts = list(width.cutoff = 60), tidy = TRUE)
options(scipen=999)
```

## Question 1

####1.1

```{r creat_vectors}
# Creating a vector to store all course numbers
courseNum <- c('MSAN501','MSAN502','MSAN504','MSAN593') 

# Creating a vector to store all course names
courseName <- c('Computation for Analytics','Review of Linear Algebra',
                'Review of Probability and Statistics','Explonatory Data Analysis') 

# Creating a vector to store names of instructors for each course
courseProf <- c('Terence J Parr','Xuemei Chen','Jeff Hamrick','Paul Intrevado')

# Creating a vector indicating which courses I formally enrolled in 
enrolled <- c(T,T,F,T)

# Creatomg a vector to store my anticipated letter grade in each course
anticipatedGrade <- c('A','B',NA,'C')

# Creating a vector to store my anticipated hours spent on each class per week
anticipatedHours <- c(20,10,NA,10)


```
Create a table summarizing the type and class for each vector. 


|Vector|Class|Type|
-------|-----|-----|
courseNum|`r class(courseNum)`|`r typeof(courseNum)`
courseName|`r class(courseName)`|`r typeof(courseName)`
courseProf|`r class(courseProf)`|`r typeof(courseProf)`
enrolled|`r class(enrolled)`|`r typeof(enrolled)`
anticipatedGrade|`r class(anticipatedGrade)`|`r typeof(anticipatedGrade)`
anticipatedHours|`r class(anticipatedHours)`|`r typeof(anticipatedHours)`

####1.2 
```{r creating dataframe}
bootcampDataFrame <- data.frame(courseNum,courseName,courseProf,
                       enrolled,anticipatedGrade,anticipatedHours)
```

|Vector|Class|Type|
-------|-----|-----|
courseNum|`r class(bootcampDataFrame$courseNum)`|`r typeof(bootcampDataFrame$courseNum)`
courseName|`r class(bootcampDataFrame$courseName)`|`r typeof(bootcampDataFrame$courseName)`
courseProf|`r class(bootcampDataFrame$courseProf)`|`r typeof(bootcampDataFrame$courseProf)`
enrolled|`r class(bootcampDataFrame$enrolled)`|`r typeof(bootcampDataFrame$enrolled)`
anticipatedGrade|`r class(bootcampDataFrame$anticipatedGrade)`|`r typeof(bootcampDataFrame$anticipatedGrade)`
anticipatedHours|`r class(bootcampDataFrame$anticipatedHours)`|`r typeof(bootcampDataFrame$anticipatedHours)`

In the data frame, some variable retain their original classes/types while some do not.

####1.3 
```{r 1.3}
bootcampDataList <- list(courseNum=courseNum,courseName=courseName,courseProf=courseProf,
                   enrolled=enrolled,anticipatedGrade=anticipatedGrade,
                   anticipatedHours=anticipatedHours)

```
|Vector|Class|Type|
-------|-----|-----|
`r names(bootcampDataList)[1]`|`r class(bootcampDataList[[1]])`|`r typeof(bootcampDataList[[1]])`
`r names(bootcampDataList)[2]`|`r class(bootcampDataList[[2]])`|`r typeof(bootcampDataList[[2]])`
`r names(bootcampDataList)[3]`|`r class(bootcampDataList[[3]])`|`r typeof(bootcampDataList[[3]])`
`r names(bootcampDataList)[4]`|`r class(bootcampDataList[[4]])`|`r typeof(bootcampDataList[[4]])`
`r names(bootcampDataList)[5]`|`r class(bootcampDataList[[5]])`|`r typeof(bootcampDataList[[5]])`
`r names(bootcampDataList)[6]`|`r class(bootcampDataList[[6]])`|`r typeof(bootcampDataList[[6]])`

In the list, all variables retain the original classes/types.

####1.4 

```{r 1.4, echo = TRUE}
# The total number of hours you anticipate spending on coursework, both per week, and over
# all of boot camp.
knitr::kable(data.frame(courseName,anticipatedHours_perweek=anticipatedHours,
                        anticipatedHours_total=anticipatedHours*5))
# A data frame with only the third row and first two columns of bootcampDataFrame
knitr::kable(bootcampDataFrame[3,1:2])

# The first value in the second element of bootcampDataList
bootcampDataList[[2]][1]
```

####1.5 

```{r 1.5}
# Convert the anticipatedGrade variable in bootcampDataFrame into an ordinal factor
bootcampDataFrame$anticipatedGrade <-       factor(bootcampDataFrame$anticipatedGrade,
                                                   levels = c('C','B','A'),ordered = T)

# The maximum letter grade you anticipate receiving in boot-camp
max(bootcampDataFrame$anticipatedGrade,na.rm = T)

# The name and course number of that class
paste(bootcampDataFrame$courseNum[which.max(bootcampDataFrame$anticipatedGrade)],
      ":",bootcampDataFrame$courseName[which.max(bootcampDataFrame$anticipatedGrade)])
```
    
## Question 2

####2.1
Read in the file `titanic.csv` and store the data in the data frame `titanicData`.
```{r 2.1}
titanicData <- read.csv("~/Desktop/titanic.csv")
```

####2.2
Rows in this data frame
```{r 2.2}
nrow(titanicData)
```

####2.3 
Columns in this data frame
```{r 2.3}
ncol(titanicData)
```

####2.4
Variable that has the most `NA` entries.
```{r 2.4}
names(titanicData)[which.max(apply(is.na(titanicData),2,FUN =sum))]
```

####2.5
Variables should be converted to a different type than the default type they were importedas.
```{r}
# Check types of variables when they were imported.
knitr::kable(sapply(titanicData,typeof),col.names = c('Type'),row.names = T)
```

`Name` should be converted to character;

`Sex` should be converted to character;

`Cabin` should be converted to character;

`Embarked` should be converted to character;

```{r}
# Convert types of those variables and store the related information in the type data frame
titanicData$Name <- as.character(titanicData$Name)
titanicData$Sex <- as.character(titanicData$Sex)
titanicData$Cabin <- as.character(titanicData$Cabin)
titanicData$Embarked <- as.character(titanicData$Embarked)

type<- data.frame(Changed_Variable=c('Name','Sex','Cabin','Embarked'),
                 DefaultType=rep('Integer',4),ChangedType=rep('Character',4))
knitr::kable(type)
```

####2.6
Coerce the `survived` variable into type `logical`
```{r}
titanicData$Survived<-as.logical(titanicData$Survived)
```

```{r}
# Mean age of survivors.
mean(titanicData$Age[titanicData$Survived==TRUE],na.rm=T)

# The mean age of those who did not survive.
mean(titanicData$Age[titanicData$Survived!=TRUE],na.rm=T)

# Plot side-by-side histograms of the ages of survivors and non-survivors.
par(mfrow=c(1,2))
hist(titanicData$Age[which(titanicData$Survived==TRUE)],
     main = "Age of Survivors",xlab = NULL)
hist(titanicData$Age[which(titanicData$Survived==FALSE)],
     main = "Age of Not Survived", xlab = NULL)
```

####2.7
```{r}
# First 10 values in cabin.
head(titanicData$Cabin,10)

#Replace all blanks with NAs.
titanicData[titanicData=='']=NA
```

####2.8
```{r}
# The percentage of the observations for age are NAs.
scales::percent(sum(is.na(titanicData$Age))/length(titanicData$Age))

# Imuptation
titanicData$Age[is.na(titanicData$Age)]=mean(titanicData$Age,na.rm = T)
```
Since `mean` could be highly influenced by some extreme values, if we fill all `NA`s with the mean, which could magnify those influence from extreme values, cause the result to be biased.

##Question 3
####3.1
Generate 100 random variables $\sim \mathcal{U}\{-1,1\}$ and compute the mean and variance (no need to set the seed for this exercise).
```{r}
r100 <- runif(100,min=-1,max=1)
mean(r100)
var(r100)
```
Repeat the previous step for sample sizes of 1,000, 10,000, 100,000 and 1,000,000, computing the mean and variance for each sample size.
```{r}
r1000 <- runif(1000,min=-1,max=1)
mean(r1000)
var(r1000)
r10000 <- runif(10000,min=-1,max=1)
mean(r10000)
var(r10000)
r100000 <- runif(100000,min=-1,max=1)
mean(r100000)
var(r100000)
r1000000 <- runif(1000000,min=-1,max=1)
mean(r1000000)
var(r1000000)
```
Create a data frame called `unifDataFrame` with seven variables: `sampleSize`, `theoreticalMean`, `sampleMean`, `deltaMean`, `theoreticalVariance`, `sampleVariance`, `deltaVariance`.
```{r}
# Create the blank data frame uniDataFrame
uniDataFrame <- data.frame(sampleSize=integer(5),theoreticalMean=numeric(5),                                  sampleMean=numeric(5),deltaMean=numeric(5),theoreticalVariance=numeric(5),
                sampleVariance=numeric(5),deltaVariance=numeric(5))
# Using a loop to generate data in uniDataFrame
for (i in (1:5)){
    uniDataFrame$sampleSize[i]=10^(i+1);
    uniDataFrame$theoreticalMean[i]=10^(i+1)*0; # 
    uniDataFrame$sampleMean[i]=mean(runif(10^(i+1),-1,1));
    uniDataFrame$deltaMean[i]=uniDataFrame$sampleMean[i]-uniDataFrame$theoreticalMean[i];
    uniDataFrame$theoreticalVariance[i]=(1+1)^2/12;
    uniDataFrame$sampleVariance[i]=var(runif(10^(i+1),-1,1));
    uniDataFrame$deltaVariance[i]=uniDataFrame$sampleVariance[i]-
        uniDataFrame$theoreticalVariance[i]
}
```

```{r}
# Plot the deltaMean.
plot(uniDataFrame$deltaMean~uniDataFrame$sampleSize)
```

```{r}
# Plot the deltaVariance.
plot(uniDataFrame$deltaVariance~uniDataFrame$sampleSize)
```

####3.2
```{r}
# Generate the vector of random variables and store them in myRunifVec
myRunifVec = runif(10000000,0,1)

# Create the histogram of the subsample
hist(sample(myRunifVec,100000))
```

####3.3
```{r}
# Create the data frame store two random variables
myRunifDataFrame <- data.frame(col1=runif(10000000,0,1),col2=runif(10000000,0,1))

# Create the third variable to sum up col1 and col2 
myRunifDataFrame$runifSum=myRunifDataFrame$col1+myRunifDataFrame$col2

# Create the histogram of runifSum to check the distribution
hist(myRunifDataFrame$runifSum)
```

####3.4
```{r}
# Create the data frame store two random variables
myRexpDataFrame <- data.frame(col1=rexp(10000000,1),col2=rexp(10000000,1))

# Create the third variable to sum up col1 and col2 
myRexpDataFrame$rexpSum=myRexpDataFrame$col1+myRexpDataFrame$col2

# Create the histogram of rexpum to check the distribution
hist(myRexpDataFrame$rexpSum)
```

##Question 4
```{r}
set.seed(100)
x_1 <- runif(100000, -100, 100)
y_1 <- rexp(100000, rate = 0.5)
```
####4.1
```{r}
# Compute the coefficients for the simple linear regression.
b1_1 <- as.numeric((x_1-mean(x_1))%*%(y_1-mean(y_1))/((x_1-mean(x_1))%*%(x_1-mean(x_1))))
b0_1 <- (sum(y_1-b1_1*x_1))/length(x_1)
```
####4.2

Since we have
  $$
  \begin{aligned}
   \hat{Y_i} = b_0 + b_1 X_i
   \\
   SSE =  \sum_{i=1}^{n} \Big(Y_i-\hat{Y}\Big)^{2}
   \\
   SSR =  \sum_{i=1}^{n} \Big(\hat{Y_i}-\bar{Y}\Big)^{2}
   \\
   SST = SSR+SSE
   \\
   R^2 = \frac{SSR}{SST}
  \end{aligned}
  $$
```{r}
# Compute SSE, SSR, SST and R2
SSE_1 <- sum((y_1-b0_1-b1_1*x_1)%*%(y_1-b0_1-b1_1*x_1))
SSR_1 <- sum((b0_1+b1_1*x_1-mean(y_1))%*%(b0_1+b1_1*x_1-mean(y_1)))
SST_1 <- sum((y_1-mean(y_1))%*%(y_1-mean(y_1)))
R2_1 <- SSR_1/SST_1
```
####4.3
```{r}
# Plot y on x, and draw the fitted regression line
plot(y_1~x_1)
abline(b0_1,b1_1)
```
##4.4
```{r}
# Compute the residuals 
e_1 <- y_1-(b0_1-b1_1*x_1)
```

##4.5
```{r}
# Generate a residual plot
plot(x_1,e_1)
abline(h=0)
```

Repeat using other distribution
```{r}
set.seed(999)
x_2 <- rnorm(100000, -100, 100)
y_2 <- rexp(100000, rate = 0.5)
```
```{r}
# Compute the coefficients for the simple linear regression.
b1_2 <- as.numeric((x_2-mean(x_2))%*%(y_2-mean(y_2))/((x_2-mean(x_2))%*%(x_2-mean(x_2))))
b0_2 <- (sum(y_2-b1_2*x_2))/length(x_2)
```
```{r}
# Compute SSE, SSR, SST and R2
SSE_2 <- sum((y_2-b0_2-b1_2*x_2)%*%(y_2-b0_2-b1_2*x_2))
SSR_2 <- sum((b0_2+b1_2*x_2-mean(y_2))%*%(b0_2+b1_2*x_2-mean(y_2)))
SST_2 <- sum((y_2-mean(y_2))%*%(y_2-mean(y_2)))
R2_2 <- SSR_2/SST_2
```
```{r}
# Plot y on x, and draw the fitted regression line
plot(y_2~x_2)
abline(b0_2,b1_2)

```

```{r}
# Compute the residuals 
e_2 <- y_2-(b0_2-b1_2*x_2)
```

```{r}
# Generate a residual plot
plot(x_2,e_2)
abline(h=0)
```

The third distribution
```{r}
set.seed(543)
x_3 <- rnorm(100000, -100, 100)
y_3 <- rnorm(100000, -100, 100)
```

```{r}
# Compute the coefficients for the simple linear regression.
b1_3 <- as.numeric((x_3-mean(x_3))%*%(y_3-mean(y_3))/((x_3-mean(x_3))%*%(x_3-mean(x_3))))
b0_3 <- (sum(y_3-b1_3*x_3))/length(x_3)
```
```{r}
# Compute SSE, SSR, SST and R3
SSE_3 <- sum((y_3-b0_3-b1_3*x_3)%*%(y_3-b0_3-b1_3*x_3))
SSR_3 <- sum((b0_3+b1_3*x_3-mean(y_3))%*%(b0_3+b1_3*x_3-mean(y_3)))
SST_3 <- sum((y_3-mean(y_3))%*%(y_3-mean(y_3)))
R2_3 <- SSR_3/SST_3
```
```{r}
# Plot y on x, and draw the fitted regression line
plot(y_3~x_3)
abline(b0_3,b1_3)

```

```{r}
# Compute the residuals 
e_3 <- y_3-(b0_3-b1_3*x_3)
```

```{r}
# Generate a residual plot
plot(x_3,e_3)
abline(h=0)
```

Create a table comparing the models
```{r}
b_0 <- c(b0_1,b0_2,b0_3)
b_1 <- c(b1_1,b1_2,b1_3)
SSE <- c(SSE_1,SSE_2,SSE_3)
SSR <- c(SSR_1,SSR_2,SSR_3)
SSTO<- c(SST_1,SST_2,SST_3)
R2 <- c(R2_1,R2_2,R2_3)
finaltable <- data.frame(b_0,b_1,SSE,SSR,SSTO,R2,
                         row.names = c('Model 1','Model 2','Model 3'))
names(finaltable) <- c("$b_0$","$b_1$","SSE","SSR","SSTO","$R^2$")
knitr::kable(t(finaltable),)
```


Since the coefficient $b_1$ and $R^2$ are really small, those x has really weak explanation power toward y, in all three models. But in the third model, since the scale of y is higher, lead to higher scale of SSE,SSR and SST, but $R^2$ actually does not improve much. 

